PhD Thesis

### 2016-11-16 main.tex
Start revising intro

\title{PhD Thesis}
\author{Paolo G. Giarrusso}
\begin{document}
\maketitle
\chapter{Intro}

A program manipulating collections of data will often perform queries on those
collections. These collections can be big, so query execution can be expensive
and warrant considerable optimization effort.
% \section{}
Efficient queries are written either as declarative database queries that are
optimized by database optimizers, or as manually optimized programs using
collection libraries. Compare maintaining anagraphic data for a population
as either a relational database or a collection:
\begin{itemize}
\item If the population data is maintained as a database, a programmer that needs to
  look up quickly people by age, can just add an index and keep the lookup query
  unchanged; the index can then be automatically maintained.
\item If the population data is maintained as an in-memory collection, adding an
  index requires modifying by hand queries to reuse it for maximum performance,
  and ensuring all changes to population data update the index to maintain
  correctness. This is error-prone and violations are common \pg{XXX}, even if
  there is quite some research on the problem.
\end{itemize}
Modern libraries offer high-level APIs for manipulating (in-memory) collections;
equivalently, such an API is an embedded domain-specific language (EDSL).
Compared to SQL, an EDSL will typically have first-class support for using
arbitrary user-defined functions in queries, because it will allow .

>>>>

\title{PhD Thesis}
\author{Paolo G. Giarrusso}
\begin{document}
\maketitle
\section*{Abstract}
% Programs manipulating data collections require

% \begin{abstract}
% \end{abstract}
\chapter{Intro}

A program manipulating collections of data will often query those collections.
For instance, a program manipulating anagraphic data about German citizens might
need to compute their average age or other anagraphic information. These
collections can be big, so query execution can be expensive and warrant
considerable optimization effort. Moreover, when input data changes, we often
want to obtain the results of a query for the new input without rerunning that
query from scratch.

% \section{}
Efficient queries are classically written either as declarative database queries that are
optimized by database optimizers, or as manually optimized code using
collection libraries. Compare maintaining anagraphic data for a population
as either a relational database or a collection:
\begin{itemize}
\item If the population data is maintained by a database management system
  (DBMS), a programmer that needs to look up quickly people by age, can just add
  an index and keep the lookup query unchanged. When the data is updated, the
  DBMS can adapt the index to the new changes, that is, it can \emph{incrementally} update the
  index automatically.
  Maintaining indexes adds possible to
  One can then experiment with different sets of indexes and workloads
\item If the population data is maintained as an in-memory collection, adding an
  index requires modifying by hand queries to reuse it for maximum performance.
  Experimenting and ensuring all changes to population data update the index to
  maintain correctness. This is error-prone, and while there is a considerable
  amount of research on this problem\pg{cite}, incorrect update code is common.
\end{itemize}
Modern libraries offer high-level APIs for manipulating (in-memory) collections;
equivalently, such an API is an embedded domain-specific language (EDSL).

Compared to SQL, EDSL queries will typically be able to use user-defined
functions in queries. Moreover, the author of EDSL queries can enjoy abstraction
mechanisms from the underlying language.
